Automated Verification for Real-Time Systems using Implicit Clocks and an Extended Antimirov Algorithm

### 2022-06-16 paper/Response.md

We acknowledge our weaknesses upon related work and evaluations. 
Thanks again for the time and effort in pointing out our typos 
and giving out constructive suggestions.

We took all the valuable comments into account and made the 
following change list.

# CHANGE LIST:

| | Changes     | Description | Timeline (Working days)    |
| ---  | ----------- | ----------- | ----------- |
|1 | Typos       | Mentioned in the reviews  | Already Done| 
|2 | Related work | [Song & Chin 2020]  | 5 days (Partly done in by responding to our Reviewers) |
|3 | Evaluation data | Have a table detailing summary statistics | 5-10 days |
|4 | Paper polishing | Find a buddy to check the writings | 5 days | 
|  | In total  | | 15 - 20 days

@@

- Drop the organization paragraph;

In total, we plan to finish all the changes within one month.

# RESPONSE:

## Common Questions Upon the Evaluation

### Q1. Where do the models come from?

The evaluation benchmark (16 selected programs) was created by ourselves. 
We developed a set of testing programs (~70 programs) during the development process.
The testing programs were designed in a functionality-oriented way, where each 
forward rules in Sec. 4.1 has 3-8 testing testing programs. The rest ~20 programs 
are mixed combinations of different constructs and some stress tests.

Our selected programs were from the last ~20 programs varying from lines of code.

### Q2. We like to explain why not choosing existing benchmarks:

Model checkers like Uppaal take timed automata directly as their inputs, 
which is not feasible for our tool.

Model checkers like PAT take timed process algebras as their inputs are more 
feasible. We indeed included some tests from their examples into our tests, 
such as the Fischer’s Mutual Exclusion Protocol (cf Fig. 4).
But they do not support value dependent bounds in their timed behavioral patterns.
However, we take the `value dependent bounds` to be one of our main novelties, therefore 
most programs in our benchmark are created ourselves, and 
have value dependent time bounds.

### Q3. What kinds of specifications you check and comparison with PAT?

All the specifications are with value-depended constructors. 
For each program, the very first specification is its strongest postcondition 
(which should be valid). Then we derived the rest specifications by 
adding various mutations (or mixture of the mutations) into the strongest postcondition, such as: 
disjunctions, negation of certain events, modification of the clock constraints, 
inserting parallel traces etc.

For the programs we included from PAT benchmark, we validate the verification 
results against PAT implementation (as in given the same logic model and the same 
specification, our tool and PAT return the same verification results). 
For the rest  specifications, we manually label the expected verification results. 
And all the results presented in Table. 3 were as expected.

### Q4. Do the authors plan to make their tool and experiments publicly available?

Yes. In fact the source code is open sourced now. We 
plan to release the benchmark after further documentation.

@@

### +P18 Is there a definition of when $\Phi \sqsubseteq \Phi’$ is valid?

Definition 2.2 (TimEffs Inclusion), at line 211.

## To Reviewer C

### Q1. Please comment on the related work I cite.

### Q2. Please comment on some of the technical questions I raised

(why talk about trace leading to a method in preconditions, 
are quantified Presburger conditions used in your benchmarks?)

Can't you not express loop invariants using specifications and reason with programs using verification conditions written for iteration-free code snippets, like in Floyd-style verification?

### Q3. Do you capture the strongest post?

Yes, it is meant to capture the strongest post.

### Q4. Are there any relative completeness results that you can show that narrow the space where your reasoning is incomplete?

### Q5. Can you give some (simple) examples where your technique does not work?

### 2022-06-16 paper/Response.md

We acknowledge our weaknesses upon related work and evaluations. 
Thanks again for the time and effort in pointing out our typos 
and giving out constructive suggestions.

We took all the valuable comments into account and made the 
following change list.

# CHANGE LIST:

| | Changes     | Description | Timeline (Working days)    |
| ---  | ----------- | ----------- | ----------- |
|1 | Typos       | Mentioned in the reviews  | Already Done| 
|2 | Related work | [Song & Chin 2020]  | 5 days (Partly done in the response) |
|3 | Evaluation data | Add more detailed explanations for the evaluation and make the benchmark and documentation public | 5-10 days |
|4 | Paper polishing | Find a buddy to check the writings | 5 days | 
|  | In total  | | 15 - 20 days

@@

#### Q2. We like to explain why not choosing existing benchmarks:

Model checkers like Uppaal take timed automata directly as their inputs, 
which is not feasible for our tool.

Model checkers like PAT take timed process algebras as their inputs are more 
feasible. We indeed included some tests from their examples into our tests, 
such as the Fischer’s Mutual Exclusion Protocol (cf Fig. 4).
But they do not support value dependent bounds in their timed behavioral patterns.
However, we take the `value dependent bounds` to be one of our main novelties, therefore 
most programs in our benchmark are created ourselves, and 
have value dependent time bounds.

#### Q3. What kinds of specifications you check and any comparisons with PAT?

@@

For the programs we included from PAT benchmark, we validate the verification 
results against PAT implementation (as in given the same logic model and the same 
specification, our tool and PAT return the same verification results). 
For the rest  specifications, we manually label the expected verification results. 
And all the results presented in Table. 3 were as expected.

#### Q4. Do the authors plan to make their tool and experiments publicly available?

Yes. In fact the source code is open sourced now. We 
plan to release the benchmark after further documentation.

---

###  To Reviewer A

Many thanks for pointing out the related works. We will be checking them 
out and add comparisons accordingly.

#### Q1. Does this mean that the time is discrete, or does this mean that the time bounds are integer-valued (while time can still be continuous)?
Time is continuous. The current implementation makes use of SMT in 
integer numbers domain.
It is not hard to adapt it into real numbers given the suitable SMT solver.

#### Q2. I find the event name Sugar slightly misleading: if this denotes the completion of the sugar addition phase, perhaps EndSugar would be more appropriate? 
`Sugar` denotes the completion of the sugar addition phase, we will change it to `EndSugar`. Thanks!

#### Q3. "our effects provide more detailed information than traditional timed verification, and in fact, it cannot be fully captured by any prior works " This requires some discussion, and does not seem obvious to me, especially when compared with Uppaal.

@@

Event `Sugar` denotes the completion of the whole sugar addition phase (which adds n portions of Sugar in total). 
We will change `Sugar` to `EndSugar` (as suggested by our review A) and revise the explanations.

#### +P2: line 90 “valid traces \Phi' … why “valid”? I thought that \Phi' is the specification of the program. Are some traces invalid? If so, where is the term "invalid trace" defined?

$\Phi^\prime$ indeed is the specification of the program. In another word, 
it is an over-approximation (or superset per se) of all the desired/expected behaviours. 
We consider any unexpected traces  are `invalid`.

#### P4 line 312, you are talking of assignments \alpha^*, but you do not define them anywhere.

@@

$\alpha$ is defined in Fig. 5 (assignment), and $\alpha^*$ is a list of $\alpha$.

#### +P5 line 318. What happens if e does not terminate on time? Does the program abort?

Like other modeling languages, the timed constructs 
are used in a abstract modeling way, where e must terminate 
before the deadline. If e is possibly go beyond the deadline, it should 
be modelled using conditionals where the other possibilities are explicitly modeled.

#### +P6 line 320 “potentially allows efficient clock manipulation”. I am not clear what you mean. Where does this clock manipulation happen?

@@

This comes from the context of how people usually make uses of the timed automata. 
For example, to use the model checker Uppaal, uses need to draw the automata 
themselves, including the clock naming, clock constrains annotation, 
transition constrains annotation. And this process most likely is not the most 
efficient way of manipulating the clocks, which leads to redundant clocks/constrains or bugs.

#### +P7 line 346, rule $[tot_1]$ should the conclusion say $(V, e_1’)$ rather than $(V, e’)$?

@@

That's right, sorry about the confusion and thanks for pointing it out. Already fixed.

#### +P8 line 360. Are we missing a rule to eliminate $\textbf{deadline}$ when the execution of $e$ hs terminated and the deadline has not been reached.

That's right. We will need to add a rule where e reduced to a value within the deadline.

#### +P9 How do you express, eg the behaviour of two consecutive delays?

@@

As you mentioned, it should be the case of: 
$(V,\textbf{delay}\ t_1; \textbf{delay}\ t_2)  \rightarrow^{t_1+t_2} (V, ())$

This is achieved by applying the sequence rules $[seq_1]$ and $[seq_2]$ and delay
rules $[delay_1]$ and $[delay_2]$ in a certain order.

#### +P10 Make sure that the definitions of $d$, $s$, $\phi$ appear before Fig. 7. The page break make mw wonder whether these were defined anywhere. Moreover, I think you should replace $\triangleq$ in line 444, with $:$. Namely $d$ is one natural number, not the set of natural numbers.

@@

Thanks! Already fixed accordingly.

#### +P11 I think the definition on line 430 is badly formed when $\pi$ is not satisfied -- ie non terminating recursion.

@@

$\textbf{A}(\nu, \alpha^*)$, $\tau (\pi)$, $\overline{A}$ and  $\_$ are possible constructors 
of $ev$ (cf Fig. 6). So the corresponding case is in line 423 for $\pi \wedge ev$.

And es is for  $\theta$, it supposed to be `\es` in our latex writing. Sorry! Fixed already.

To define the LHS of 431, it includes all the RHS from lines 431 - 434.

#### +P14 line 470, I think you also need to say that $\{\pi_i,\theta_i\}\in\{\Pi,\Theta\}$

@@

#### +P15 Lines 505-509. I think the definition does not take into consideration that it is possible for $e_1$ to execute and apply some side-effects to the $V$, then for $d$ to be exceeded before $e_1$ terminates, and then we continue with $e_2$. I think the rule needs to consider the side-effects of the incomplete $e_1$ too.

There is no possible incompletion of $e_1$ in timeout. 
Informally, in process $e_1$ timeout[𝑑] $e_2$, the first 
observable event of
$e_1$ shall occur before d time units elapse 
(since the process starts). Otherwise, $e_2$ takes over control
after exactly 𝑑 time units elapse.

So it's either $e_1$ or $e_2$ to be executed. The formal semantics can 
be found in $[to_1]$ to $[to_4]$ (on top of the page 8).

#### +P18 Is there a definition of when $\Phi \sqsubseteq \Phi’$ is valid?

@@

Definition 2.2 (TimEffs Inclusion), at line 211.

---

###  To Reviewer C

#### Q1. Please comment on the related work I cite.

Specification of real-time and hybrid systems in rewriting logic, by Olveczky and Meseguer.

Specification and Analysis of Real-Time and Hybrid Systems in Rewriting Logic - Olveczky (PhD thesis) 
See also: Verification of clocked and hybrid systems- Kesten, Manna, Pnueli

You are also missing references to work done by Cook and Koskinen and others on LTL/CTL verification of programs, which also include reductions to standard Hoare-style verification.

#### Q2. Please comment on some of the technical questions I raised

1. I find it strange that you don't have iterative loops. Why is that?

The language allows each iterative
loop to be optimized to an equivalent tail-recursive method, where mutation
on parameters is made visible to the caller. The technique of translating away
iterative loops is standard and is helpful in further minimising our core language.

2. Why talk about trace leading to a method in preconditions?

We politely disagree upon the opinion mentioned in the review: 
`While verification of a method does depend on the values of variables at the time it's called, the events that happened earlier seem irrelevant.`

Just like particular requirements of the input values, there are possible 
requirements upon the history traces. 
For example, before reading a file, we wish to make sure the file was actually opened 
(and not yet closed) is the history. This can be written in precondition as:
$\Phi_{pre} = \_^\star {\cdot} Open {\cdot} ({\overline{Close}})^\star $. 
Such examples extend to the timed processes in different contexts.

3. Are quantified Presburger conditions used in your benchmarks?

We do not have quantified Presburger conditions in current benchmarks. 
Indeed 'Purely universal quantification' better describes our implementation. 
(Although having TRS solving quantified Presburger conditions with traces seems 
useful and challenging.)

#### Q3. Do you capture the strongest post?

To be precise, it only captures the 'strongest over-approximation'.

The forward verifier creates over-approximation when there are conditionals 
(cf [FV-If-Else-Local] on top of page 11).

For example, $\Phi_{post}{=} (n {>} 0 \wedge A{\cdot}B)  \vee (n {\leq} 0 \wedge C{\cdot}D)$

Therefore in theorem 4.1,
if we have $\{P\}\ e\ \{Q\}$ 
 we can only say there exists a post in Q which captures the 
actual execution.

#### Q4. Are there any relative completeness results that you can show that narrow the space where your reasoning is incomplete?
#### Q5. Can you give some (simple) examples where your technique does not work?

We like to answer Q4 and Q5 together.

For example, while proving this following (expected to be valid) inclusion: 
$ t_1 {<}3 {\wedge} t_2 {<}4 {\wedge}  (\epsilon \# t_1) {\cdot} (\epsilon \# t_2) 
\sqsubseteq  t {<}7 {\wedge} (\epsilon \# t)$

The current rewriting would simple unify $t_1$ and $t$ leaving $t_2$ behind, which 
leads to the next step:

$ t_2 {<}4 {\wedge}  (\epsilon \# t_2) 
\sqsubseteq   \epsilon $, then eliminate $(\epsilon \# t_2) $ from the both side, 
leads to the next step:

$True {\wedge} \epsilon \sqsubseteq   \bot$, then disproves this inclusion.

This can be solved by adding an axiom: $(\epsilon \# t_1) {\cdot} (\epsilon \# t_2) = 
(\epsilon \# (t_1{+}t_2))$. 
However, we did not include it because it seems to be leading to other incomplete situations.

#### Q6. Am I reading it correctly that your system is able to prove all valid properties and disprove all invalid ones?
Yes. cf  Q3 in `Common Questions Upon the Evaluation`.

### 2022-07-06 paper/tv.tex
m

%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
\documentclass[acmsmall,10pt,review]{acmart}

@@

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract} 
The correctness of real-time systems 
depends both on the correct functionalities and the realtime constraints.
To go beyond the existing Timed Automata based techniques, 
we propose a novel solution that integrates a 
modular Hoare-style forward verifier with a new term rewriting 
system (TRS) on \emph{Timed Effects} (\timedEffects).
The main purposes are to increase the expressiveness,  
dynamically create clocks, 
and efficiently solve constraints on the clocks.  
We formally define 
a core language \timedL, generalizing the real-time systems, modeled 
using mutable variables and timed behavioral patterns, 
such as \emph{delay}, \emph{timeout}, \emph{deadline}. 
Secondly, to capture real-time specifications, 
we introduce \timedEffects, a new effects logic, 
that extends 
\emph{Regular Expressions} with dependent
values and arithmetic constraints.
Thirdly,  the forward verifier infers temporal behaviors of given 
\timedL\ programs, expressed in \timedEffects. 
Lastly, we present a purely algebraic TRS, i.e., an extended \emph{Antimirov algorithm}, to 
efficiently prove language inclusions between 
 \timedEffects. 
To demonstrate the feasibility of our proposals, 
we prototype the verification system; prove its 
soundness; report on case studies and the experimental results.

@@

    \m{(Program)} &  \code{\mathcal{P} }&\code{  ::= }&
    \code{%{datat^*}  \  
   (\alpha^*, {meth^*})} 
%\qquad   \qquad\quad 
 %   \m{(Basic\ Types)} \qquad  \code{b} \code{\quad  ::= \quad }
  %   \code{int\ |\ bool\ |\ void}
    \\
    \m{(Assignment)} &  \code{\alpha }&\code{  ::= }&
    \code{x:=v} 
   \\

%     \m{(Data\ Struct.)} &  \code{datat}&\code{  ::= }&
 %    \code{\textbf{struct}\ d\ \{ {(t\ f)^*} \}}  
 %   \\

%     \m{(Types)} &  \code{t}&\code{  ::= }&
% \code{d\ |\ \tau} 
% \\    
  \m{(Method\ Def.)} &  \code{meth}&\code{  ::= }&
  \code{ mn  \ {(x)^*}\ \{\textbf{req}\ \effect_{pre}\ \textbf{ens}\  \effect_{post}\} \ \{ e \}}
  \\
  \m{(Values)} & \code{v} &\code{  ::= } & ()\ |\ c \ | \ b \ |\ x\  
  \\
  \m{(Expressions)} &  \code{e}&\code{  ::= }&
    \code{ v \ | \ 
    [b]\ e \ |\ mn({v^*}) \ }
  \code{| 
 % v.f:=e \ | 
 \ e_1;e_2\ | \ e_1 || e_2 \ 
  |\ {if}\ b\ {then}\ e_1\ {else}\ e_2 \
 }  \\

 &   & & \code{
  |\ \textbf{event} [\anyevent{A}(v, \alpha^*)] 
  | \ \textbf{delay} [d] \ 
  | \ e_1\ \textbf{timeout} [d] \  e_2 \ 
  | \ \textbf{deadline}  [d] \ e \  
  }  \\
  ~\\

    \multicolumn{4}{c}{
 % \code{k^\tau:constant\ of\ type\ \tau\ } 
 % \qquad\quad
\code{c  \in \mathbb{Z}}
\qquad \ 
\code{b  \in \mathbb{B}}
\qquad \  
\code{x, mn  ::\in \textbf{var}}  
\qquad\ 
 \code{(Action\ labels)\ \anyevent{A}  \in \Sigma %\cup \{\tau\}
\qquad\ 
\m{(Time\ Bounds)}\ \code{d} \code{\in \mathbb{N}}    
}   
    }\\  
    \hline
  \end{array}  
  $
 \caption{A core first-order imperative language with timed constructs, \timedL.} 
 \label{fig:code_language}

@@

    \m{(Pure)}  & \code{\code{{\pi}}}&\code{  ::= }&
{ {True}}
   \ |  {False}
   \ |  {A(}{t_1, t_2}{)}
  \  |   {{\pi_1}}  \wedge  \code{{\pi}}_2
   \ |  {{\pi_1}} {\vee} \code{{\pi}}_2
   \ |  \neg\code{{\pi}}
  \  |   {{\pi_1}} {\Rightarrow} {{\pi_2}}
   \ |  \forall x. \code{{\pi}}
   \ |  \exists x. \code{{\pi}}

@@

    \m{(Real\text{-}Time\ Term)}  & \code{t} &\code{  ::= }&
    \code{ c
    \ | \  x
    \ | \  t_1{+}t_2
    \ | \  t_1\text{-}t_2
    %\ | \ k {\times} n
   }

    \\~\\
    \multicolumn{4}{c}{
     % \code{(Events)  \anyevent{A} ::\in \Sigma}   
   %\qquad
    \code{c ::\in \mathbb{Z}} 
     \qquad \qquad 
           \code{x  ::\in \textbf{var} } 
           \qquad\qquad

@@

The  syntax of the \timedEffects\ is formally defined in 
\figref{fig:Syntax_of_Types_and_Effects} Effects is 
 a conditioned event sequence \code{{\pi}\wedge es} or a disjunction of 
 two effects $\effect_1 \vee \effect_2$.
Timed sequences comprise \textit{nil} ($\bot $);
 an empty trace $ \epsilon$;
a single event \code{ev};
  %where \anyevent{a} is an element from a finite set of events, denoted by \code{\Sigma};
%parametrized event \code{ \anyevent{a}.\iota} where \code{\iota} is a list pattern expressing a possibly infinite list (for example, \code{\anyevent{a}.[1,2,3,4]},  \code{\anyevent{a}.[1 .. 4]} and \code{\anyevent{a}.[1, 2.. 4]} all refer to the finite event sequence \code{\anyevent{a}.1 \cdot \anyevent{a}.2 \cdot \anyevent{a}.3 \cdot \anyevent{a}.4},
%while \code{\anyevent{a}.[1 ..]} refers to the infinite event sequence \code{\anyevent{a}.1 \cdot \anyevent{a}.2 \cdot \anyevent{a}.3 \cdot}..., and \code{\anyevent{a}.[1, 3..]} refers to the infinite event sequence \code{\anyevent{a}.1 \cdot \anyevent{a}.3 \cdot \anyevent{a}.5\cdot}...);
 sequences concatenation \code{\es_1\cdot \es_2};
disjunction  \code{\es_1\vee \es_2};
parallel composition of two processes is written as \code{\es_1 || \es_2}, where \code{\es_1} and \code{\es_2} 
may communicate via multi-party event synchronization or shared variables;
a blocking waiting for a certain constraint to be satisfied \code{\pi ?}; 
We introduce a new operator \code{\mysharp }, and the effects 
\code{\es \mysharp t} represents that a trace takes real-time 
\code{t} to complete, where \code{t} is a \emph{term}. 
A timed sequence can be constructed by \code{\es^\star}, representing zero or 
more times repetition of the trace \code{\es}.

@@

{{\small\begin{gather*}
  \code{\frac{
    \begin{matrix}
      \vdash \{\Pi, \epsilon \}\ e_1 \ \{  \Pi_1, \Theta_1\} 
      \qquad 
      \vdash \{ \Pi, \epsilon \}\ e_2 \ \{  \Pi_2, \Theta_2 \} 
      \\
      \s{\Pi_f, \Theta_f} {=} \s{\Pi_1  \wedge t_1{\leq}d, (hd(\Theta_1)\mysharp t \cdot tl(\Theta_1))}
      \cup 
  \s{\Pi_2  \wedge t_2{>}d,  (\Theta_2\mysharp t)}   
  \quad (t_1, t_2\ are\ fresh)     
\end{matrix}
  }
  {\vdash \s{ \Pi, \Theta} \  
  e_1\ \textbf{timeout} [d] \  e_2
   \ \s{\Pi_f, \Theta \cdot \Theta_f} } \   {[FV\text{-}Timeout]}} 
\end{gather*}}}

>>>>

%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
\documentclass[acmsmall,10pt,review,anonymous]{acmart}

@@

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract} 
The correctness of real-time systems 
depends both on the correct functionalities and the realtime constraints.
To go beyond the existing Timed Automata based techniques, 
we propose a novel solution that integrates a 
modular Hoare-style forward verifier with a new term rewriting 
system (TRS) on \emph{Timed Effects} (\timedEffects).
The main purposes are to increase the expressiveness,  
dynamically create clocks, 
and efficiently solve constraints on the clocks.  
We formally define 
a core language \timedL, generalizing the real-time systems, modeled 
using mutable variables and timed behavioral patterns, 
such as \emph{delay}, \emph{timeout}, \emph{interrupt}, \emph{deadline}. 
Secondly, to capture real-time specifications, 
we introduce \timedEffects, a new effects logic, 
that extends 
\emph{Regular Expressions} with dependent
values and arithmetic constraints.
Thirdly,  the forward verifier infers temporal behaviors of given 
\timedL\ programs, expressed in \timedEffects. 
Lastly, we present a purely algebraic TRS, i.e., an extended \emph{Antimirov algorithm}, to 
efficiently prove language inclusions between 
 \timedEffects. 
To demonstrate the feasibility of our proposals, 
we prototype the verification system; prove its 
soundness; report on case studies and the experimental results.

@@

    \m{(Program)} &  \code{\mathcal{P} }&\code{  ::= }&
    \code{%{datat^*}  \  
   (\alpha^*, {meth^*})} 
    \\

    \m{(Basic\ Types)} &   \code{\iota} &\code{  ::=  }&
     \code{int\ |\ bool\ |\ void}

    \\
%     \m{(Data\ Struct.)} &  \code{datat}&\code{  ::= }&
 %    \code{\textbf{struct}\ d\ \{ {(t\ f)^*} \}}  
 %   \\

  \m{(Method\ Def.)} &  \code{meth}&\code{ ::= }&
  \code{\iota\  mn  \ {(\iota\ x)^*}\ \{\textbf{req}\ \effect_{pre}\ \textbf{ens}\  \effect_{post}\} \ \{ e \}}
  \\
  \m{(Values)} & \code{v} &\code{  ::= } & ()\ |\ c \ | \ b \ |\ x\  
  \\

  \m{(Assignment)} &  \code{\alpha }&\code{  ::= }&
  \code{x:=v} 
 \\
  \m{(Expressions)} &  \code{e}&\code{  ::= }&
    \code{ v \ 
    | \ \alpha \ | \ 
    [b]\ e \ |\ mn({v^*}) \ }
  \code{|  \ e_1;e_2\ | \ e_1 || e_2 \ 
  |\ {if}\ b\ {then}\ e_1\ {else}\ e_2 \
 }  \\

 &   & & \code{
  |\ \textbf{event} [\anyevent{A}(v, \alpha^*)] \ 
  | \ \textbf{delay} [t] \ 
  | \ e_1\ \textbf{interrupt} [t] \  e_2 \ 
  }  \\

  &   & & \code{
    | \ e_1\ \textbf{timeout} [t] \  e_2 \ 
    | \ e \ \textbf{deadline}  [t] \ 
    | \ e \ \textbf{waituntil} [t]  \ 
    }  \\

    \m{(Real\text{-}Time\ Terms)}  & \code{t} &\code{  ::= }&
    \code{ c
    \ | \  x
    \ | \  t_1{+}t_2
    \ | \  t_1\text{-}t_2
   }

    \\
    ~\\

    \multicolumn{4}{c}{
 % \code{k^\tau:constant\ of\ type\ \tau\ } 
 % \qquad\quad
\code{c  \in \mathbb{Z}}
\qquad \qquad 
\code{b  \in \mathbb{B}}
\qquad \qquad  
\code{x, mn  \in \textbf{var}}  
\qquad \qquad 
 \code{(Action\ labels)\ \anyevent{A}  \in \Sigma %\cup \{\tau\}
%\qquad\ 
%\m{(Time\ Bounds)}\ \code{d} \code{\in \mathbb{N}}    
}   
    }\\  
    \hline
  \end{array}  
  $
 \caption{A core first-order imperative language with timed constructs, \timedL.} 
 \label{fig:code_language}

@@

    \m{(Pure)}  & \code{\code{{\pi}}}&\code{  ::= }&
{ {True}}
   \ |  {False}
   \ |  {A(}{t_1, t_2}{)}
  \  |   {{\pi_1}}  \wedge  \code{{\pi}}_2
   \ |  {{\pi_1}} {\vee} \code{{\pi}}_2
   \ |  \neg\code{{\pi}}
  \  |   {{\pi_1}} {\Rightarrow} {{\pi_2}}
  % \ |  \forall x. \code{{\pi}}
  %s \ |  \exists x. \code{{\pi}}

@@

    \m{(Real\text{-}Time\ Terms)}  & \code{t} &\code{  ::= }&
    \code{ c
    \ | \  x
    \ | \  t_1{+}t_2
    \ | \  t_1\text{-}t_2
    %\ | \ k {\times} n
   }

    \\~\\
    \multicolumn{4}{c}{
     % \code{(Events)  \anyevent{A} ::\in \Sigma}   
   %\qquad
    \code{c \in \mathbb{Z}} 
     \qquad \qquad 
           \code{x  \in \textbf{var} } 
           \qquad\qquad

@@

The  syntax of the \timedEffects\ is formally defined in 
\figref{fig:Syntax_of_Types_and_Effects} Effects is 
 a conditioned event sequence \code{{\pi}\wedge es} or a disjunction of 
 two effects $\effect_1 \vee \effect_2$.
Timed sequences comprise \textit{nil} ($\bot $);
 an empty trace $ \epsilon$;
a single event \code{ev};
 sequences concatenation \code{\es_1\cdot \es_2};
disjunction  \code{\es_1\vee \es_2};
parallel composition of two processes is written as \code{\es_1 || \es_2}, where \code{\es_1} and \code{\es_2} 
may communicate via multi-party event synchronization or shared variables;
a blocking waiting for a certain constraint to be satisfied \code{\pi ?}; 
We introduce a new operator \code{\mysharp }, and the effects 
\code{\es \mysharp t} represents that a trace takes real-time 
\code{t} to complete, where \code{t} is a \emph{term}. 
A timed sequence can be constructed by \code{\es^\star}, representing zero or 
more times repetition of the trace \code{\es}.

@@

{{\small\begin{gather*}
  \code{\frac{
    \begin{matrix}
      \vdash \{\Pi, \epsilon \}\ e_1 \ \{  \Pi_1, \Theta_1\} 
      \qquad 
      \vdash \{ \Pi, \epsilon \}\ e_2 \ \{  \Pi_2, \Theta_2 \} 
      \\
      \s{\Pi_f, \Theta_f} {=} \s{\Pi_1  \wedge t_1{\leq}d, (hd(\Theta_1)\mysharp t_1 \cdot tl(\Theta_1))}
      \cup 
  \s{\Pi_2  \wedge t_2{>}d,  (\Theta_2\mysharp t)}   
  \quad (t_1, t_2\ are\ fresh)     
\end{matrix}
  }
  {\vdash \s{ \Pi, \Theta} \  
  e_1\ \textbf{timeout} [d] \  e_2
   \ \s{\Pi_f, \Theta \cdot \Theta_f} } \   {[FV\text{-}Timeout]}} 
\end{gather*}}}
